---
title: "R & Elasticsearch: elastic package"
author: "Scott Chamberlain"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    theme: yeti
---


```{r echo=FALSE}
hook_output <- knitr::knit_hooks$get("output")
knitr::knit_hooks$set(output = function(x, options) {
  lines <- options$output.lines
  if (is.null(lines)) {
    return(hook_output(x, options))  # pass to default hook
  }
  x <- unlist(strsplit(x, "\n"))
  more <- "..."
  if (length(lines) == 1) {        # first n lines
    if (length(x) > lines) {
      # truncate the output, but add ....
      x <- c(head(x, lines), more)
    }
  } else {
    x <- c(if (abs(lines[1])>1) more else NULL,
           x[lines],
           if (length(x)>lines[abs(length(lines))]) more else NULL
    )
  }
  # paste these lines together
  x <- paste(c(x, ""), collapse = "\n")
  hook_output(x, options)
})

knitr::opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  comment = "#>"
)
```


```{r}
library(elastic)
```

## Connect

```{r}
connect()
```

connect with complete error messages

```{r}
connect(errors = "complete")
```

```{r eval=FALSE}
Search(index = "asdfasdfsdfsa")
#> Error: 404 - no such index
#> ES stack trace:
#>
#>   type: index_not_found_exception
#>   reason: no such index
#>   resource.type: index_or_alias
#>   resource.id: asdfasdfsdfsa
#>   index: asdfasdfsdfsa
```

## Ping, is server up

```{r}
ping()
```

## Indices

```{r}
cat_indices()
```

```{r}
index_exists("shakespeare")
```

```{r}
index_delete("shakespeare")
```

```{r}
shakespeare <- system.file("examples", "shakespeare_data.json", package = "elastic")
invisible(
  docs_bulk(shakespeare)
)
```

## Search

```{r}
res <- Search(index = "shakespeare")
```

## Access results

took, timed_out, shards

```{r}
res$took
res$timed_out
res$`_shards`
```

hits

```{r}
res$hits
```

total hits, max score

```{r}
res$hits$total
res$hits$max_score
```

## Parse hits

Most users will want to get to a data.frame

```{r}
library('data.table')
hits <- res$hits$hits
src <- lapply(hits, "[[", "_source")
data.table::rbindlist(src)
```

> It would be hard to parse results for the user - data structure unpredictable

## URI Search

E.g., when you can't pass a body, perhaps `POST` requests are blocked

```{r}
### wildcards, either * or ?
Search_uri(index="shakespeare", q="*ay")$hits$total

### fuzziness
Search_uri(index="shakespeare", q="text_entry:ma~")$hits$total

### Proximity searches
Search_uri(index="shakespeare", q='text_entry:"as hath"~5')$hits$total
```

## Back to Search

Fields

```{r}
Search(index="shakespeare", fields=c('play_name','speaker'))$hits$hits[[1]]
```

Sorting

```{r}
Search(index="shakespeare", type="act", sort="text_entry")$hits$hits[1:2]
```

```{r}
Search(index="shakespeare", type="act", sort="speaker:desc", fields='speaker')$hits$hits[1:2]
```

Paging

```{r}
Search(index="shakespeare", size=1, from=1, fields='text_entry')$hits
```

Queries

```{r}
Search(index="shakespeare", type="act", q="speaker:KING HENRY IV")$hits$total
```

## Search body

```{r}
aggs <- '{
    "aggs": {
        "latbuckets" : {
           "histogram" : {
               "field" : "decimalLatitude",
               "interval" : 5
           }
        }
    }
}'
res <- Search(index="gbif", body=aggs, size=0)
res <- res$aggregations$latbuckets$buckets
df <- data.table::setDF(data.table::rbindlist(res))
library(ggplot2)
ggplot(df, aes(key, doc_count)) + geom_line(size = 3) + theme_grey(20)
```

## curl options

Common options are `verbose()`, `timeout()`, `progress()`, `config(followlocation=TRUE)`.

```{r}
library('httr')
out <- Search(config = verbose())
out <- Search(config = progress())
```
