Examples working with data on the web in R
=====

## Interactive data in R

```{r rcharts1}
library(rCharts)

# Load a data set
data(economics, package = "ggplot2")

# Make the date column of date format
econ <- transform(economics, date = as.character(date))

# Make a javascript plot object
m1 <- mPlot(x = "date", y = c("psavert", "uempmed"), type = "Line", data = econ)

# Set attributes
m1$set(pointSize = 0, lineWidth = 1)

# Visualize
m1$show(cdn=TRUE)
```

```{r rcharts2}
# Load a data set
hair_eye_male <- subset(as.data.frame(HairEyeColor), Sex == "Male")

# Make a javascript plot object
n1 <- nPlot(Freq ~ Hair, group = "Eye", data = hair_eye_male, type = "multiBarChart")

# Visualize
n1$show(cdn=TRUE)
```

```{r rcharts3}
# Make a new map object
map3 <- Leaflet$new()

# Set view dimensions and zoom
map3$setView(c(51.505, -0.09), zoom = 13)

# Add two markers
map3$marker(c(51.5, -0.09), bindPopup = "<p> Hi. I am a popup </p>")
map3$marker(c(51.495, -0.083), bindPopup = "<p> Hi. I am another popup </p>")

# Visualize
map3$show(cdn=TRUE)
```


## Politicians talk - SunlightFoundation listens

### Look at mentions of the phrase "climate change" in congress

```{r govdat}
library(govdat)
library(ggplot2)

# Get mentions of climate change from Democrats
dat_d <- sll_cw_timeseries(phrase='climate change', party="D")

# Add a column that says this is data from deomcrats
dat_d$party <- rep("D", nrow(dat_d))

# Get mentions of climate change from Democrats
dat_r <- sll_cw_timeseries(phrase='climate change', party="R")

# Add a column that says this is data from republicans
dat_r$party <- rep("R", nrow(dat_r))

# Put two tables together
dat_both <- rbind(dat_d, dat_r)

# Plot data
ggplot(dat_both, aes(day, count, colour=party)) + geom_line() + 
  theme_grey(base_size=20) + scale_colour_manual(values=c("blue","red"))
```


## NOAA climate data, using the rnoaa package

### Poking around things  you can do with rnoaa

```{r pokearound}
library(rnoaa)

# Search for hourly precipitation data in zip code 28801 on a single day
noaa(dataset='PRECIP_HLY', location='ZIP:28801', datatype='HPCP', year=1980, month=7, day=17)$data

# Search sations including the word "snake"
noaa_search(dataset='GHCND', resulttype='station', text='snake')

# Search countries including the word "boulder"
noaa_search(dataset='GHCND', resulttype='city', text='boulder')
```


### Example 1

#### Search for normal daily max temperature and plot it!

```{r rnooa1_, comment=NA, warning=FALSE, message=FALSE, cache=TRUE}
library(rnoaa)
library(scales)
library(ggplot2)

# Search for NOAA data 
out <- noaa(dataset='NORMAL_DLY', station='GHCND:USW00014895', datatype='dly-tmax-normal', year=2010, month=4)

# Plot data
noaa_plot(out)

# Where is this station
temp <- noaa_stations(dataset='NORMAL_DLY', station='GHCND:USW00014895')
latlong <- data.frame(temp$data$meta[c('latitude','longitude')])
world <- map_data(map='world', region='.')
ggplot(world, aes(long, lat)) + # make the plot
  geom_polygon(aes(group=group), fill="white", color="gray40", size=0.2) +
  geom_point(data=latlong, aes(longitude, latitude), size=6)
```


### Example 2, sea ice!

#### Map a single year/month/pole combo

```{r seaice1}
# Get URLs for data
urls <- seaiceeurls(mo='Apr', pole='N', yr=1990)

# Create directory to store data in
dir.create("seaicedata")

# Download sea ice data
out <- noaa_seaice(url=urls, storepath="seaicedata")

# Plot data
ggplot(out, aes(long, lat, group=group)) + 
  geom_polygon(fill="steelblue") +
  theme_ice()
```

#### Map all years for April only for North pole

```{r seaice2}
library(doMC)
library(plyr)

# Get URLs for data
urls <- seaiceeurls(mo='Apr', pole='N')[1:12]

# Download sea ice data
registerDoMC(cores=4)
out <- llply(urls, noaa_seaice, storepath="seaicedata", .parallel=TRUE)

# Name elements of list
names(out) <- seq(1979,1990,1)

# Make a data.frame
df <- ldply(out)

# Plot data
ggplot(df, aes(long, lat, group=group)) + 
  geom_polygon(fill="steelblue") +
  theme_ice() +
  facet_wrap(~ .id)
```


### Example 3, a bit more complicated

#### Load some packages

```{r rnoaa_load, comment=NA, warning=FALSE, message=FALSE, cache=TRUE}
library(lubridate)
library(doMC)
library(plyr)
library(stringr)
```

#### Find stations to get data from

```{r findstations, comment=NA, warning=FALSE, message=FALSE, cache=TRUE}
# Get list of stations
stations <- noaa_stations(dataset='GHCND', enddate='20121201')

# Get station IDs
res <- sapply(stations$data, function(x) x$meta$id)
```

#### Get data from stations, just searching on the first 60.

```{r noaa_stations, comment=NA, warning=FALSE, message=FALSE, cache=TRUE}
# Make the function noaa work even when data not available
noaa_fw <- failwith(NULL,noaa)

# Search for data
registerDoMC(cores=4)
dat <- compact(llply(as.character(res)[1:20], function(x) noaa_fw(dataset='GHCND', station=x, year=2010, month=7), .parallel=TRUE))
```

#### Make a data.frame and fix dates. 

```{r noaa_getdata, comment=NA, warning=FALSE, message=FALSE, cache=TRUE}
df <- ldply(dat, function(x) x$data)
df$date <- ymd(str_replace(as.character(df$date), "T00:00:00\\.000", ''))
df <- df[df$dataType == 'PRCP',]
```

#### Plot precipitation by day for each station

```{r noaa_plot, comment=NA, warning=FALSE, message=FALSE, cache=TRUE}
library(ggplot2)

ggplot(df, aes(date, value)) +
  theme_bw(base_size=18) + 
  geom_line(size=2) +
  facet_grid(station~., scales="free") +
  scale_x_datetime(breaks = date_breaks("7 days"), labels = date_format('%d/%m/%y')) +
  labs(y=as.character(df[1,'dataType']), x="Date") +
  noaa_theme() +
  theme(strip.text.y = element_text(angle=0))
```




## World Bank climate data, using the rWBclimate package

### Example 1: look at increases in temperature over time.

It is possible to extract historical data from GCM queries, but this is acutally model backcasting output, not true historical records. The climate data api can download data at the same spatial scales as all other requests (country or basin), but the time scale differs. You can download data at monthly, yearly or decadanal time scales. Monthly data is actually the mean monthly temperature or precipitation value for each month from 1901-2009 for countries, or 1960-2009 for basins. You can indicate the type of data you want in the call by setting the time_scale parameter to month, *year* or decade.

```{r rwbclimate1, comment=NA, warning=FALSE, message=FALSE, cache=TRUE}
library(rWBclimate)

countries <- c("USA", "MEX", "CAN", "BLZ")
dat <- get_historical_temp(countries, "year")

ggplot(dat, aes(x = year, y = data, group = locator)) + 
  geom_point() + 
  geom_path() + 
  labs(y="Average annual temperature", x="Year") +
  theme_bw(base_size=22) +
  stat_smooth(se = F, colour = "black") + 
  facet_wrap(~locator, scale = "free")
```


### Example 2: Plotting annual data for different countries

Data can be extracted from countries or basins submitted as vectors. Here we will plot the expected temperature anomaly for each 20 year period over a baseline control period of 1961-2000. These countries chosen span the north to south pole. It's clear from the plot that the northern most countries (US and Canada) have the biggest anomaly, and Belize, the most equatorial country, has the smallest anomaly.

```{r}
# Search for data
country.list <- c("CAN", "USA", "MEX", "BLZ", "ARG")
country.dat <- get_model_temp(country.list, "annualanom", 2010, 2100)

# Subset data to one specific model
country.dat.bcc <- country.dat[country.dat$gcm == "bccr_bcm2_0", ]

# Exclude A2 scenario
country.dat.bcc <- subset(country.dat.bcc, country.dat.bcc$scenario != "a2")

# Plot data
ggplot(country.dat.bcc, aes(x = fromYear, y = data, group = locator, colour = locator)) + 
  geom_point() + 
  geom_path() + 
  ylab("Temperature anomaly over baseline") + 
  theme_bw(base_size=20)
```



## Phenology data using rnpn (interacts with the US National Phenology Network)

```{r rnpn1}
library(rnpn)

# Lookup names
temp <- lookup_names(name='bird', type='common')
comnames <- temp[temp$species_id %in% c(357, 359, 1108), 'common_name']

# Get some data
out <- getobsspbyday(speciesid=c(357, 359, 1108), startdate='2010-04-01', enddate='2013-09-31')
names(out) <- comnames
df <- ldply(out)
df$date <- as.Date(df$date)

# Visualize data
library(ggplot2)
ggplot(df, aes(date, count)) + 
  geom_line() +
  theme_grey(base_size=20) +
  facet_grid(.id ~.)
```