`ro warning=FALSE, message=FALSE, comment=NA, cache=FALSE or`

*********

## Example rOpenSci code

*********

### Getting data - ropensnp (wraps the OpenSNP.org API)
```{r ropensnp}
# install_github("ropensnp", "ropensci")

library(ropensnp)
dat <- phenotypes(userid = "1,6,8", df = TRUE)
dat

```

*********

### Literature  - metadata - rdatacite (wraps the DataCite API)
```{r rdatacite}
# install_github('rdatacite', 'ropensci')

library(rdatacite); library(OAIHarvester)
out <- getrecord("56225", T) # get the metadata record with id '56225'
out$metadata # get $identifier, $datestamp, $setSpec, or $metadata
oaih_transform(out$metadata) # transform only metadata to a list
```

*********

### Literature - full text - rplos (wraps the Public Library of Science API)
```{r rplos}
# install_github("rplos", "ropensci")

library(rplos)

# Visualize with the ggplot2 package, only up to two terms for now
plot_throughtime(list('drosophila','monkey'), 100)

# Search results on a keyword over all fields in PLoS Journals.
plosword('Helianthus')
out <- plosword(list('monkey','replication','design','sunflower','whale'), vis = 'TRUE')
out[[1]] # results in a data frame 
out[[2]] # results in a bar plot
```

*********

### Altmetrics - raltmet (wraps many APIs)
```{r raltmet}
# install_github("raltmet", "ropensci")

library(raltmet)

# Get github stats on repos
github(userorg = 'hadley', repo = 'ggplot2', 'allstats')

# Get Bit.ly clicks for a given URL.
bitly_clicks(bitlyurl = 'http://bit.ly/xaTrO7')
```

*********

### NCBI data - rentrez (wraps the entrez NCBI API)
```{r rentrez}
# install_github("rentrez", "ropensci")

library(rentrez)

# Get summaries of objects in NCBI datasets from a unique ID
pubmed_search <- entrez_search(db="pubmed", term="Dwarf Elephant", retmax=1)
entrez_summary(db="pubmed", ids=pubmed_search$ids)
```

*********

### Get the .Rmd file used to create this post. You can run the code just as you would in R. You can also compile to pdf, html, etc. using [knitr](http://yihui.name/knitr/). For example, to compile to pdf, run `knit2pdf("/path/to/theRmdfile.Rmd")`.