---
title       : Communities, networks, & visualizations in R
date        : 2013-07-31
author      : Scott Chamberlain
job         : 
framework   : io2012        # {io2012, html5slides, shower, dzslides, ...}
highlighter : highlight.js  # {highlight.js, prettify, highlight}
hitheme     : tomorrow      # 
widgets     : []            # {mathjax, quiz, bootstrap}
mode        : selfcontained # {standalone, draft}
assets      :
  css: "http://netdna.bootstrapcdn.com/font-awesome/3.2.1/css/font-awesome.css"
---

## Where to find this presentation
<br><br><br>
### The presentation = https://bitly.com/sfuworkshop
<br><br><br>
### The code (written in R Markdown) = http://bit.ly/1bKVX2O

---

## The plan

+ Data manipulation
+ General visualizations
+ Networks/graphs: analyses and viz
+ Community ecology: analyses and viz

---

<!-- Data manipulation -->

## Data manipulation

<br><br><br><br>

## plyr + reshape2 = <i class="icon-smile"></i>

---

## Intro to subsetting
<br><br><br><br>
### See here [https://github.com/hadley/devtools/wiki/Subsetting](https://github.com/hadley/devtools/wiki/Subsetting)

---

## A few quick examples

```{r manip1, cache=TRUE, comment=NA}
library(reshape2)
head(iris)[1:3,]
```

```{r manip2, cache=TRUE, comment=NA}  
iris_m <- melt(iris, id.vars=5)
head(iris_m)[1:3,1:3]
```

```{r manip3, cache=TRUE, comment=NA, warning=FALSE, message=FALSE}
dcast(iris_m, Species ~ variable, fun.aggregate=sd)
```

---

## What can we do with this?

`manipulate, then plot`

```{r manipplot1, cache=TRUE, fig.height=6, fig.width=11, tidy=FALSE, fig.align='center'}
iris_m <- melt(iris, id.vars=5)
ggplot(iris_m, aes(Species, value)) +
  geom_boxplot() +
  facet_wrap(~ variable) +
  theme_grey(base_size=20)
```

---

## Your turn

<br><br><br><br>

### Read in a data file, melt the data, then cast the data, applying a function to that data.

---

## plyr - split, apply, combine

### Function naming scheme: <font color="#B38249">first letter of source R object</font> + <font color="#E09E38">first letter of output R object</font> + ply

+ data.frame = ddply, dlply, daply
+ list = llply, ldply, laply
+ array/vector: aaply, alply, adply
+ matrix: maply , mlply, mdply

---

```{r plyr1, cache=TRUE, comment=NA}
data <- data.frame(x = c("a", "a", "b", "b", "c", "c"), y = c(2, 4, 0, 5, 5, 10))
ddply(data, .(x), summarise, mean(y))
```

```{r plyr2, cache=TRUE, comment=NA}
data <- data.frame(x = c("a", "a", "b", "b", "c", "c"), y = c(2, 4, 0, 5, 5, 10))
data_list <- list(data, data, data)
ldply(data_list, function(z) ddply(z, .(x), summarise, mean(y)))
```

---

## DIY plyr

<br><br><br><br>

### Perform split-apply-combine for an R object of your choice. 


<!-- Visualizations -->

---

## ggplot2 terminology

+ ggplot - The main function where you specify the dataset and variables to plot
+ geom - geometric objects
  + `geom_point()`, `geom_bar()`, `geom_density()`, `geom_line()`, `geom_area()`
+ aes - aesthetics
  + shape, alpha (transparency), color, fill, linetype
+ scale - Define how your data will be plotted
  + continuous, discrete, log

---

```{r viz1, eval=FALSE, echo=TRUE}
install.packages("ggplot2")
```

```{r viz2, cache=TRUE, fig.height=5.5, tidy=FALSE, fig.align='center'}
library(ggplot2)
ggplot(data=iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point() + 
  theme_grey(base_size=20)
```

---

## Building blocks, mix and match

```{r buildup1, eval=FALSE, cache=TRUE, tidy=FALSE, fig.align='center'}
ggplot(data=iris, aes(Sepal.Length, Sepal.Width, colour=Species)) +
  geom_point() + 
  theme_grey(base_size=20)
```

OR, just do 

```{r buildup2, eval=FALSE, cache=TRUE, tidy=FALSE, fig.align='center'}
p <- ggplot(data=iris, aes(Sepal.Length, Sepal.Width, colour=Species))
p <- p + geom_point()
p + theme_grey(base_size=20)
```

---

## Color by species

```{r color1, cache=TRUE, fig.height=5.5, tidy=FALSE, fig.align='center'}
ggplot(data=iris, aes(Sepal.Length, Sepal.Width, colour=Species)) +
  geom_point() +
  theme_grey(base_size=20)
```

--- &twocol

## Adjust the size (and color) of points

*** =left

```{r size1, cache=TRUE, fig.height=5.5, tidy=FALSE, fig.align='center'}
ggplot(data=iris, aes(Sepal.Length, Sepal.Width, colour=Species)) +
  geom_point(size = 6) +
  theme_grey(base_size=20)
```

*** =right

```{r size2, cache=TRUE, fig.height=5.5, tidy=FALSE, fig.align='center'}
ggplot(data=iris, aes(Sepal.Length, Sepal.Width, color=Species, size=Species)) +
  geom_point() +
  theme_grey(base_size=20)
```

---

## Facet by species

```{r viz4, cache=TRUE, fig.height=5.5, fig.width=11, tidy=FALSE, fig.align='center'}
ggplot(data=iris, aes(Sepal.Length, Sepal.Width)) +
  geom_point() +
  facet_wrap(~ Species) +
  theme_grey(base_size=20)
```

--- &twocol

## Combine geoms

*** =left

### Combine `geom_boxplot` and `geom_point`

```{r combinegeoms1, cache=TRUE, fig.height=5.5, tidy=FALSE, fig.align='center'}
ggplot(data=iris, aes(Species, Petal.Width)) +
  geom_boxplot() +
  geom_point()
```

*** =right

### Order matters! 

```{r combinegeoms2, cache=TRUE, fig.height=5.5, tidy=FALSE, fig.align='center'}
ggplot(data=iris, aes(Species, Petal.Width)) +
  geom_point() +
  geom_boxplot()
```

---

## Make this plot

```{r makethis, cache=TRUE, echo=FALSE, fig.height=7, fig.width=10, tidy=FALSE, fig.align='center', message=FALSE}
ggplot(data=iris, aes(Petal.Width, fill=Species)) +
  geom_histogram() +
  facet_wrap(~ Species) +
  theme_grey(base_size=20)
```

--- 

## Saving plots using ggplot2

If the plot is on your screen
```{r saveggplot1, eval = FALSE}
ggsave('~/path/to/figure/filename.png')
```

If your plot is assigned to an object
```{r saveggplot2, eval = FALSE}
ggsave(plot1, file = "~/path/to/figure/filename.png")
```

Specify a size
```{r saveggplot3, eval = FALSE}
ggsave(file = "/path/to/figure/filename.png", width = 6,
height =4)
```

Or any format (pdf, png, eps, svg, jpg)
```{r saveggplot4, eval = FALSE}
ggsave(file = "/path/to/figure/filename.jpg")
ggsave(file = "/path/to/figure/filename.pdf")
```

<!-- networks -->
<!-- networks -->
<!-- networks -->

--- 

## Networks

+ Visualizations
+ Analyses
  + Network level metrics
  + Species level metrics

---

## Network vizualizations

```{r netviz1, cache=TRUE, comment=NA, fig.height=8, fig.width=14, fig.align='center'}
library(bipartite)
plotweb(small1976)
```

---

## Another way to visualize networks

```{r netviz2, cache=TRUE, fig.height=8, fig.width=14, fig.align='center'}
visweb(small1976, labsize=2)
```

---

## Variety of other plotting options

```{r netviz3, eval=FALSE}
# Eg 1
plotweb(small1976, high.lablength=3, low.lablength=0, arrow="down")

# Eg 2
plotweb(small1976, text.rot=90, arrow="down.center", col.interaction="wheat2",
  y.lim=c(-1,2.5))

# Eg 3
low.abun = round(runif(dim(small1976)[1],1,40))
names(low.abun) <- rownames(small1976)
plotweb(small1976, text.rot=90, low.abun=low.abun, col.interaction="purple", 
  y.width.low=0.05, y.width.high=0.05)
```

---

## Network metrics - of the whole community

Start with the bipartite package. Others include X, Y, and Z.

```{r netlevelmets1, cache=TRUE, comment=NA, fig.height=5.5, fig.align='center'}
res <- networklevel(small1976, index="binary")
res
```

---

## Let's compare some metrics among networks!!!

```{r comparemets, cache=TRUE, comment=NA, warning=FALSE, fig.height=3.5, fig.width=10, fig.align='center'}
library(reshape2)
networks <- list(Safariland,barrett1987,bezerra2009,elberling1999,inouye1988,kato1990,kevan1970)
res <- llply(networks, function(x) networklevel(x, index=c('connectance','links per species','nestedness')))
names(res) <- c('Safariland','barrett1987','bezerra2009','elberling1999','inouye1988','kato1990','kevan1970')
df <- melt(ldply(res), id.vars='.id')
ggplot(df, aes(.id, value)) +
  geom_point(size=4, alpha=0.6) + 
  facet_wrap(~variable, scales="free") +
  theme_bw(base_size=20) +
  theme(axis.text.x = element_blank())
```

---

## Make this plot

```{r comparemets2, echo=FALSE, cache=TRUE, comment=NA, warning=FALSE, fig.height=7, fig.width=12, fig.align='center'}
networks <- list(Safariland,barrett1987,bezerra2009,elberling1999,inouye1988,kato1990,kevan1970)
res <- llply(networks, function(x) networklevel(x, index=c('web asymmetry','H2')))
names(res) <- c('Safariland','barrett1987','bezerra2009','elberling1999','inouye1988','kato1990','kevan1970')
df <- melt(ldply(res), id.vars='.id')
ggplot(df, aes(.id, value, color=.id)) +
  geom_point(size=4, alpha=0.6) + 
  facet_wrap(~variable, scales="free") +
  theme_bw(base_size=20) +
  theme(axis.text.x = element_blank())
```

---

## Species level metrics

```{r splevelmets, cache=TRUE, comment=NA, warning=FALSE}
splevel <- specieslevel(small1976, index="degree")
head(splevel[[1]], n=3)
head(splevel[[2]], n=3)
```

---

## Species level metrics - roll your own

+ Pick a network in the bipartite package
+ Calculate one species level metric for all species in that network
+ Plot the metrics for each species, both plants and pollinators

<!-- community structure -->
<!-- community structure -->
<!-- community structure -->

--- 

## Community structure

+ Diversity indices
+ Rarefaction - comparing diverity in different samples
+ Ordination

--- &twocol align1:left

## Diversity indices

*** =left

Shannon-Weaver

```{r diversity1, comment=NA, cache=TRUE}
library(vegan)
data(BCI)
bci_subset <- BCI[1:3,]
head(bci_subset)[,10:11]
(H <- diversity(bci_subset))
```

*** =right

Simpson

```{r diversity2, comment=NA, cache=TRUE}
diversity(bci_subset, index="simpson")
```

Evenness

```{r diversity4, comment=NA, cache=TRUE}
H <- diversity(bci_subset)
H/log(specnumber(bci_subset))
```

---

## Rarefaction

```{r rarefaction1, cache=TRUE, comment=NA, fig.height=5.5, fig.align='center'}
nosp <- specnumber(BCI)
nosp_rare <- rarefy(BCI, raremax)
df <- data.frame(nosp, nosp_rare)
head(df)
```

---

## Make this plot

```{r rarefaction2, echo=FALSE, cache=TRUE, comment=NA, fig.height=7, fig.width=10, fig.align='center', warning=FALSE}
ggplot(df, aes(nosp, nosp_rare)) +
  geom_point(size=4) +
  theme_grey(base_size=20) +
  scale_x_continuous(limits=c(70,110)) +
  scale_y_continuous(limits=c(70,110))
```

--- &twocol

## Rarefaction curves

*** =left 

```{r rarefaction3, cache=TRUE, comment=NA, fig.height=5.5, fig.align='center'}
raremax <- min(rowSums(BCI))
rarecurve(BCI, step=20, sample=raremax, col="blue", cex=0.6)
```

*** =right

```{r rarefaction4, cache=TRUE, comment=NA, fig.height=5.5, fig.align='center', warning=FALSE}
data(BCI)
sp1 <- specaccum(BCI)
sp2 <- specaccum(BCI, "random")
plot(sp1, ci.type="poly", col="blue", lwd=2, ci.lty=0, ci.col="lightblue")
boxplot(sp2, col="yellow", add=TRUE, pch="+")
```

---

## Ordination

```{r ordinationbase1, cache=TRUE, comment=NA, message=FALSE, warning=FALSE}
data(dune)
ord <- metaMDS(dune, trace=0)
plot(ord)
```

---

## In ggplot2!

```{r ordinationggplot, cache=TRUE, comment=NA, fig.height=5.5, fig.align='center'}
ord_axes <- data.frame(scores(ord))
ord_spp <- data.frame(scores(ord, display="species"))
ggplot(ord_axes, aes(NMDS1, NMDS2)) +
  geom_point(shape=21) +
  geom_point(data=ord_spp, aes(NMDS1, NMDS2), shape=3, colour="red")
```

---

## But I want species names

```{r ordinationggplot2, cache=TRUE, comment=NA, fig.height=5.5, fig.align='center'}
ord_spp$spp <- row.names(ord_spp)
ggplot(ord_axes, aes(NMDS1, NMDS2)) +
  geom_point(shape=21) +
  geom_text(data=ord_spp, aes(NMDS1, NMDS2, label=spp), colour="red")
```

--- &twocol

## Comparison between treatments

*** =left

```{r adonis1, cache=TRUE, comment=NA}
data(dune)
data(dune.env)
dune[1:3,1:3]
dune.env[1:3,1:3]
help(dune.env)
```

*** =right

```{r adonis2, cache=TRUE, comment=NA}
data.frame(adonis(dune ~ Management*A1, data=dune.env, permutations=99)$aov.tab)
```

---

## Do your own ordination

+ Take one of the datasets in the vegan pacage
+ Pick an ordination method
  + `cca`: constrained ordination
  + add environmental variables to ordination, see `envfit`
  + `adonis`: Permutational Multivariate Analysis of Variance
+ Try it out!

--- 

## Resources

+ All this code, etc. are available [here](http://bit.ly/1bKVX2O), and as slides [here](https://bitly.com/sfuworkshop)
+ You can poke around on the [The R homepage](http://cran.r-project.org/)
+ [R task views](http://cran.r-project.org/web/views/)
+ [Nice R documentation site](http://www.rdocumentation.org/)
+ [R mailing lists](http://www.r-project.org/mail.html), including general, stats, etc., and **ecology**!
+ [Search the R tag on StackOverflow (aka SO)](http://stackoverflow.com/questions/tagged/r): best place to ask questions, fastest response
+ Googling R? Use **"cran"** instead of **"R"** (cran=comprehensive R archive network): often R help list or on SO